Computing exact restricted isometry constants via mixed integer semidefinite programming

### 2014-12-18 RIP-MISDP-Paper/Paper_v1.tex
added macros, replaces \mathbb{R} by \R (from dsfonts)

\documentclass{elsarticle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\newtheorem{thm}{Theorem}
\newdefinition{rmk}{Definition}

@@

Finding the sparsest solution of an underdetermined system of linear equations is a problem that has been the subject of much recent research in compressed sensing. Given a matrix $A \in \mathbb{R}^{m \times n}$ with $m < n$ and 
right-hand side $b \in \mathbb{R}^m$, we are interested in a solution of 

@@

\begin{rmk} \label{ripdef}
The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \mathbb{R}^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k := 
\{x \in \mathbb{R}^n : \|x\|_0 \leq k \}$ we have
 \begin{equation}\label{rip}
  (1-\delta_k)\|x\|_2^2 \leq \|Ax\|_2^2 \leq (1+\delta_k)\|x\|_2^2.
 \end{equation}
 \end{rmk}

@@

\begin{thm} \label{RIPsqrt2}
 If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \mathbb{R}^n$
 \begin{equation*}\label{l1ErrorRIP}
  \|x^* - x \|_{\ell_1} \leq C \|x - x_k \|_{\ell_1},
 \end{equation*}
where $x^*$ is an optimal solution for (\ref{l1min}) with right-hand side $b = Ax$, $x_k$ is the best approximation of $x$ with $k$-sparse vectors and $C$ is a constant depending only on $k$ and $\delta_{2k}$. If $x$ is $k$-sparse, then the right-hand side of
(\ref{l1ErrorRIP}) is zero, so $x$ is the only $k$-sparse vector with $A\hat{x} = Ax$ and recovered by (\ref{l1min}) and therefore the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.\end{thm}

@@

\begin{thm}\label{Rk1thm}
 For every $A \in \mathbb{R}^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\textbf{Rk}(X^*) = 1$. 
\end{thm}
\begin{proof}
 Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) (because of $\textbf{Tr}(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1, z_1 = 1, z_i = 0 \ \forall i \neq 0$ is always feasible)
 with $\sum_{i=1}^n z_i = \ell \leq k$. Define
 \begin{align}\label{Proj}
 \text{min} \quad & \textbf{Tr}(\tilde{A}^T\tilde{A}X) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \tag{Proj} \\
 & X \succeq 0 , X \in S_\ell, \nonumber
 \end{align}
where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those columns $\{{i_1}, {i_2}, ..., {i_\ell}\} =: T$ of $A$ for which $\hat{z}_{i_k} = 1$. \\
Now let $\tilde{X}$ be the matrix consisting of those rows and columns 
$i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$. Then $\hat{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$ 
and $(A^TA)\hat{X}$, which do not influence the trace or the positive semidefiniteness. \\
On the other hand we can lift any optimal solution $\tilde{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
and $\bar{X}_{i_j i_k} = \tilde{X}_{jk}$ if and only if $i_j, i_k \in T$ and $\bar{X}_{ij}=0$ otherwise. \\
So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
$\textbf{Rk}(\bar{X}) = \textbf{Rk}(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\mathbb{R}^{\ell(\ell+1)/2}$ such a point exists as a consequence of corollary 18.5.1 in
\cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
\begin{equation*}
 \frac{1}{2}\textbf{Rk}(\breve{X})\cdot(\textbf{Rk}(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
\end{equation*}
As $\breve{X}=0$ is infeasible for $\textbf{Tr}(X)=1$, we must have $\textbf{Rk}(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
\end{proof}

>>>>

\documentclass{elsarticle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{dsfont}
\newtheorem{thm}{Theorem}
\newdefinition{rmk}{Definition}

\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\supp}{supp}

\newcommand{\suchthat}{\,:\,}
\newcommand{\abs}[1]{\lvert{#1}\rvert}
\newcommand{\card}[1]{\lvert{#1}\rvert}
\newcommand{\define}{\coloneqq}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\Norm}[2]{\lVert{#1}\rVert_{#2}}
\newcommand{\skal}[2]{\langle{#1},{#2}\rangle}
\newcommand{\order}[1]{O\left({#1}\right)}
\newcommand{\T}{^{\top}}

% sets
\newcommand{\R}{\mathds{R}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\ones}{\mathds{1}}

@@

Finding the sparsest solution of an underdetermined system of linear equations is a problem that has been the subject of much recent research in compressed sensing. Given a matrix $A \in \R^{m \times n}$ with $m < n$ and 
right-hand side $b \in \R^m$, we are interested in a solution of 

@@

\begin{rmk} \label{ripdef}
The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \R^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k := 
\{x \in \R^n : \|x\|_0 \leq k \}$ we have
 \begin{equation}\label{rip}
  (1-\delta_k)\|x\|_2^2 \leq \|Ax\|_2^2 \leq (1+\delta_k)\|x\|_2^2.
 \end{equation}
 \end{rmk}

@@

\begin{thm} \label{RIPsqrt2}
 If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \R^n$
 \begin{equation*}\label{l1ErrorRIP}
  \|x^* - x \|_{\ell_1} \leq C \|x - x_k \|_{\ell_1},
 \end{equation*}
where $x^*$ is an optimal solution for (\ref{l1min}) with right-hand side $b = Ax$, $x_k$ is the best approximation of $x$ with $k$-sparse vectors and $C$ is a constant depending only on $k$ and $\delta_{2k}$. If $x$ is $k$-sparse, then the right-hand side of
(\ref{l1ErrorRIP}) is zero, so $x$ is the only $k$-sparse vector with $A\hat{x} = Ax$ and recovered by (\ref{l1min}) and therefore the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.\end{thm}

@@

\begin{thm}\label{Rk1thm}
 For every $A \in \R^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\textbf{Rk}(X^*) = 1$. 
\end{thm}
\begin{proof}
 Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) (because of $\textbf{Tr}(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1, z_1 = 1, z_i = 0 \ \forall i \neq 0$ is always feasible)
 with $\sum_{i=1}^n z_i = \ell \leq k$. Define
 \begin{align}\label{Proj}
 \text{min} \quad & \textbf{Tr}(\tilde{A}^T\tilde{A}X) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \tag{Proj} \\
 & X \succeq 0 , X \in S_\ell, \nonumber
 \end{align}
where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those columns $\{{i_1}, {i_2}, ..., {i_\ell}\} =: T$ of $A$ for which $\hat{z}_{i_k} = 1$. \\
Now let $\tilde{X}$ be the matrix consisting of those rows and columns 
$i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$. Then $\hat{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$ 
and $(A^TA)\hat{X}$, which do not influence the trace or the positive semidefiniteness. \\
On the other hand we can lift any optimal solution $\tilde{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
and $\bar{X}_{i_j i_k} = \tilde{X}_{jk}$ if and only if $i_j, i_k \in T$ and $\bar{X}_{ij}=0$ otherwise. \\
So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
$\textbf{Rk}(\bar{X}) = \textbf{Rk}(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$ such a point exists as a consequence of corollary 18.5.1 in
\cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
\begin{equation*}
 \frac{1}{2}\textbf{Rk}(\breve{X})\cdot(\textbf{Rk}(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
\end{equation*}
As $\breve{X}=0$ is infeasible for $\textbf{Tr}(X)=1$, we must have $\textbf{Rk}(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
\end{proof}
